---
title: "Final Report: Your Title Here"
author: "Don Francisco"
date: "December 5, 2018"
output:
    github_document: default
bibliography: references.bib
csl: bioinformatics.csl
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

Add about 2-3 pages here. Across the whole manuscript, you should cite at least 20 peer reviewed articles.

# Methods

## Study design

Add about half a page here. In this section instead of first person (I/we), use the authors of the paper you selected, since you'll just be describing what they did, based on the methods in their paper.

## Sample origin and sequencing

Add about half a page here. In this section instead of first person (I/we), use the authors of the paper you selected, since you'll just be describing what they did, based on the methods in their paper.

## Computational

These are the methods you used to do your bioinformatic analyses analyses. Should probably be between 0.5 and 1 pages. At a very minimum should include citations for DADA2 and phyloseq if you are doing an amplicon study, or other citations as appropriate.

# Results

## Subsections are ok in the results section too

```{r your-code-here}
# Add code chunks as needed for your analyses
# For most analyses, I would recommend splitting the
# intensive computational part into a seperate R script file
# and then just load your libraries and the data object here
# using the `load()` function. Ask me for clarification if this is
# unclear.
```

```{r load-packages-and-data}

# load general-use packages
library("dplyr")
library("tidyr")
library("knitr")
library("ggplot2")

# this package allows for the easy inclusion of literature citations in our Rmd
# more info here: https://github.com/crsh/citr
# and here:
# http://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html
library("citr")

# These are the primary packages well use to clean and analyze the data
# this package needs to be installed from bioconductor -- it's not on CRAN
# see info here: https://benjjneb.github.io/dada2/dada-installation.html
library("dada2")

# This to export a fasta of our final denoised sequence variants
library("seqinr")

# To install this you have to install from GitHub
# See more info here: https://github.com/leffj/mctoolsr
# run this -- install.packages("devtools")
# and then this -- devtools::install_github("leffj/mctoolsr")
library("mctoolsr")

# And this to visualize our results
# it also needs to be installed from bioconductor
library("phyloseq")

# load data produced from analysis scripts using
# something like load("output/processed_data.Rdata")
```


```{r extract-sample-and-file-names}
# NOTE: Much of the following follows the DADA2 tutorials available here:
# https://benjjneb.github.io/dada2/tutorial.html
# Accessed October 19, 2017
# set the base path for our input data files
path <- "data/raw_data"

# Sort ensures samples are in order
filenames_forward_reads <- sort(list.files(path, pattern = ".fastq"))

# Extract sample names, assuming filenames have format: SAMPLENAME.fastq
sample_names <- sapply(strsplit(filenames_forward_reads, "\\."), `[`, 1)

# Specify the full path to each of the filenames_forward_reads
filenames_forward_reads <- file.path(path, filenames_forward_reads)
```





In addition to a minimum of 5-8 figures/tables (and associated captions), you should include sufficient text in this section to describe what your findings were. Remember that in the results section you just describe what you found, but you don't interpret it - that happens in the discussion. 2-3 pages.

# Discussion

Add around 3-4 pages interpreting your results and considering future directions one might take in analyzing these data.

# Sources Cited
